### 2017-08-06 manuscript.tex
Initial commit of the template

% use paper, or submit
% use 11 pt (preferred), 12 pt, or 10 pt only

%\documentclass[letterpaper, preprint, paper,11pt]{AAS}	% for preprint proceedings
\documentclass[letterpaper, paper,11pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,12pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, paper,10pt]{AAS}		% for final proceedings (20-page limit)
%\documentclass[letterpaper, submit]{AAS}			% to submit to JAS

\usepackage{AAS_packages}
%\usepackage{subfigure} % have subcaption in use instead
%\usepackage[notref,notcite]{showkeys}  % use this to temporarily show labels

\PaperNumber{XX-XXX}

\begin{document}

\title{Title}

\author{Shankar Kulumani\thanks{Doctoral Student, Mechanical and Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052, Email: \href{mailto:skulumani@gwu.edu}{skulumani@gwu.edu}.},  
Taeyoung Lee\thanks{Associate Professor, Mechanical and Aerospace Engineering, George Washington University, 800 22nd St NW, Washington, DC 20052, Tel: 202-994-8710, Email: \href{mailto:tylee@gwu.edu}{tylee@gwu.edu}.}
}

\maketitle{}

\begin{abstract}
Abstract
\end{abstract}

\section{Introduction}\label{sec:introduction}
% MOTIVATION FOR LOW THRUST ORBITAL TRANSFER

% use \citenum for AAS papers

\section{System Model}\label{sec:pcrtbp}

\subsection{Jacobi Integral}

\section{Variational Integrator}\label{sec:discrete_var}

\subsection{Variational Principle}

\subsection{Discrete Variational Mechanics}

\subsection{Discrete Equations of Motion}

\section{Invariant Manifolds} \label{sec:invariant_manifold}

\section{Conclusions}\label{sec:conclusion}

\bibliographystyle{AAS_publication} 
\bibliography{library}

\appendix
\section*{Appendix A: Costate Equations of Motion}\label{sec:costate_appendix}

\end{document}